Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
capturing

\subsection{Automatic activity sensing and digital well-being}
The recent proliferation in personal wearable activity sensors and
network and storage- enabled digital health measurement devices has
been driven by both the falling costs of embeddable electronics, and
the demand for self-measurement from the ``quantified-self''\cite{}
movement.  Wireless sensors such as the Fit-bit\cite{} and Nike
FuelBand\cite{} and small, discreet, and low-cost, providing the
wearer with high-resolution temporal data about their activity levels,
with little or no effort.

In addition to fitness monitors, the market for quantified self sensors has
expanded into areas of personal health. For example, there are sensors for
measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.

>>>>

The recent proliferation in personal wearable activity sensors and
network and storage- enabled digital health measurement devices has
been driven by both the falling costs of embeddable electronics, and a
surging public interest in ``quantified-self''-style self-monitoring
for self-improvement.  Many of these sensors are lightweight,
can be worn discreetly, measure several physiological statistics
simultaneously, and most require little maintenance. Table
\ref{sensors} lists a variety of such sensors, currently available
readily from online retailers.

\begin{table*}[htb]
\small
\begin{tabular}{p{5.0cm}  p{4cm} l}

 measurement & products & price range \\
\hline
general activity (walking/jogging) & FitBit, Nike FuelBand  & 49-199 \\ 
weight & Writhings WiFi Scale & 200 \\
blood pressure & Writhings BPM & 200 \\ 
sleep duration & Zeo Pro, WakeMate, & 100-300  \\
heart rate variability & Zensorium Tinke & 200 \\
blood oxygenation & Zensorium Tinke & 200 \\ 
respiratory rate & Zensorium Tinke & 200 \\ 
galvanic skin repsonse & BodyMedia CORE 2 &  50-100 \\
body temperature & BodyMedia CORE 2 &  50-100 \\ 
Eating & Hapifork & (not yet released) \\ 
\end{tabular}
\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
\end{table*}

While few of these devices had been independently evaluated for
accuracy for clinical use, as they have matured, several of these
manufacturers begun rigorous evaluaton of their products for achieving
certification. For example, BodyMedia advertise themselves as
providing the ``FDA-approved, clincally validated, most accurate
wireless activity measurement''.  A separate study of activity-based
sleep monitoring tools revealed that ``low cost actigraphy-based
approaches'' correlated well with baseline (in-lab EEG-based
measurements) for sleep duration, although less with sleep quality
metrics.

%% In addition to fitness monitors, the market for quantified self sensors has
%% expanded into areas of personal health. For example, there are sensors for
%% measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
%% Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
%% Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
%% The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
